What to simulate? Inferring the right direction for mental rotation

### 2014-04-30 man/cogsci-2014/cogsci-2014.tex
More tweaks, to fit on 6 pages

\addtolength{\abovecaptionskip}{-0.25cm}
\addtolength{\belowcaptionskip}{-0.35cm}

@@

The only models (of which the authors are aware) that seriously
attempted to address the decision of \textit{what} to simulate are
those by \citeA{Funt:1983wn} and \citeA{Just:1985uu}. In both of these
models, the axis and direction of rotation are computed prior to
performing the rotation. One object is then rotated through the target
rotation, and is checked against the other object for
congruency. However, this approach assumes that the corresponding
points on the two objects can be easily identified, which is not
necessarily the case.  Indeed, the state-of-the-art in computer vision
suggests that there is more to this problem than checking for
congruency, particularly when the shapes are complex or not exactly
the same \cite<e.g.,>{Belongie:2002tj,Sebastian:2003vm}. Additionally,
recent research shows that when performing \textit{physical}
rotations, people do not rotate until congruency is reached; they may
even rotate \textit{away} from initially near perfect matches
\cite{Gardony:2013gn}.

If people are not computing the rotation beforehand, what might they
be doing? To answer this question, we perform a rational analysis of
the problem of mental rotation
\cite{Marr:1983to,anderson90,Shepard:1987tt}. At the computational
level, we can say that the \textit{problem} is to determine which
spatial transformations an object has undergone based on two images of
that object (which do not include information about point
correspondences). At the algorithmic level, we are constrained by the
notion that mental images must be transformed in an analog manner (or
in a way that is at least approximately analog), and that mental
images are time-consuming and effortful to generate. Thus, the
\textit{goal} is to make this determination while performing a minimum
amount of computation (i.e., as few rotations as possible).

The original ``congruency'' hypothesis \cite{Shepard1971} actually is
a rational solution to this problem, in the sense that the smallest
amount of computation coincides with rotating through the minimum
angle. However, it violates the constraint that we do not know the
points of correspondence between the images, which is exactly what
necessitates the use of imagery. Noting that a rational solution need
not necessarily maintain a single trajectory of rotation, we explore
an alternative model, which---rather than explicitly computing the
angle of rotation---engages in an \textit{active sampling} strategy.

Active sampling is the idea people gather new information in a manner
that increases certainty about the problem space. An everyday example
of this can be observed in the game of ``20 questions'', in which one
person thinks of a concept, and another has to guess the concept in 20
questions or less. The first question is almost always ``person,
place, or thing?'', because the answer provides the most possible
information about the concept of interest. Active sampling has
recently gained support across several areas of cognitive science
\cite<e.g.,>{Gureckis:2012gu}, including other spatial domains
\cite{Juni:2011vo}. In the case of mental rotation, actively choosing
rotations may be the best way to gather evidence about the similarity
between the observed shapes when the angle of rotation is unknown.

@@

@@

The task we are interested in modeling involves observing two images
and determining whether one image depicts the ``same'' object as the
other image (differing only by a rotation), or a ``flipped'' version
of the object in the other image (differing by a reflection and then a
rotation).

Formally, we denote the shapes as $X_a$ and $X_b$ and assume $X_b$ is
generated by a transformation of $X_a$, i.e. $X_b=f(X_a, \theta, h)$,
where $\theta$ is a rotation, $\hi$ is the hypothesis that the images
depict the same object, and $\hf$ is the hypothesis that the images
depict mirror-image objects. The posterior probability of each
hypothesis given the observed shapes is then: $p(h\ \vert\ X_a, X_b)
\propto \int p(X_b\ \vert\ X_a, \theta, h)p(h)p(\theta)\dif\theta$,
where $p(X_b\ \vert\ X_a, \theta, h)$ is the probability that $X_b$
was generated from $X_a$. Because we ultimately want to determine
which hypothesis is more likely, the quantity of interest is a ratio
$\ell:=p(\hi\ \vert\ X_a, X_b) / p(\hf\ \vert\ X_a, X_b)$ which
(assuming that all rotations are equally likely) is equivalent to:
\begin{equation}
  \ell = \frac{\left(\int p(X_b\ \vert\ X_a, \theta, \hi)\dif\theta\right)\cdot{}p_0}{\left(\int p(X_b\ \vert\ X_a, \theta, \hf)\dif\theta\right)\cdot{}p_1},
  \label{eq:lh-ratio}
\end{equation}
where $p_0=p(h=0)$ and $p_1=p(h=1)$, for brevity. If $\ell > 1$, then
we accept the hypothesis that the images depict the same object
($\hi$); if $\ell < 1$, then we accept the hypothesis that the images
depict flipped objects ($\hf$).

@@

Additionally, we assume that the observed shapes can only be
transformed by a small amount at a time, and each transformation takes
a non-negligible amount of time. Specifically, if the current mental
image is $\Xt$, then:
\begin{equation}
  \mathbf{X}_{t+1} = \left\{ \begin{array}{ll}
      f(\Xt, 0, \epsilon) &\mbox{ rotate by $\epsilon$ radians,} \\
      f(\Xt, 1, 0) &\mbox{ flip,} \\
      f(\Xa, 0, 0) &\mbox{ reset to $0^\circ$, or} \\
      f(\Xa, 1, 0) &\mbox{ reset and flip,} \\
    \end{array} \right.
  \label{eq:actions}
\end{equation}
where $\epsilon\sim \left|\mathcal{N}(0, \sigma_\epsilon^2)\right|$
and $\sigma_\epsilon^2$ is the variance of the step size.

@@

If we give equal weight to the two hypotheses, then the priors cancel
out; if we weigh one hypothesis more heavily, then our decision will
be biased towards that hypothesis. However, unless the likelihood
ratio is already very close to 1, small biases in the prior will not
make much of a difference as to whether we accept $\hi$ or $\hf$.

@@

A model which does not know point correspondences could use the
following algorithm: (1) pick a random direction; (2) take a single
step; (3) if that step decreased similarity, then begin rotating in
the reverse direction, otherwise continue rotating in the original
direction; (4) continue rotating in the chosen direction until a
``match'' is found (defined as finding a value of $S$ that exceeds a
threshold); and (5) if no match was found, flip, and start over from
step one.  We assume that the locations where $S$ is greater than the
threshold correspond to the true $\theta$ (or points near the true
$\theta$). So, as with the \Oc{} model, we use
Equation~\ref{eq:mle-lh-ratio}. Additionally, we only allow for the
``flip'' action after no match has been found, because there is no
particularly principled way for the \Th{} model to choose when to
flip.

@@

In the current formulation of the problem, choosing the threshold is
straightforward because we know both the exact geometry of the shapes
and that a linear transformation exists which will perfectly align
them. However, this choice is not always clear \textit{a priori}, as
the global optimum depends on many factors (e.g., shape complexity,
dimensionality, perceptual uncertainty, and whether the shapes are
identical).  One way to deal with the problem of choosing a threshold
would simply be to perform an exhaustive search or use a global
optimization strategy. However, these strategies would not result in
the linear RT found by \citeA{Shepard1971}. A second alternative is to
perform a Hill Climbing (\Hc{}) search; i.e., rotate in the direction
that increases similarity until no further improvement can be
found. In contrast with the \Th{} model, this results in arriving in a
\textit{local} maximum (which may or may not be the global
maximum). Thus, as with the \Oc{} and \Th{} models, we use
Equation~\ref{eq:mle-lh-ratio}. We only allow for the ``flip'' action
after a local maximum has been reached, because like the \Th{} model,
there is otherwise no principled way for the \Hc{} model to choose
when to flip.

@@

While the previous few models all focused on \textit{searching} for
the global maximum, we need only approximate
Equation~\ref{eq:lh-ratio} with a set of sequential rotations. Based
on evidence that people use \textit{active sampling}
\cite<e.g.,>{Gureckis:2012gu}, we hypothesize a model that actively
monitors the usefulness of the rotations it is performing. Instead of
searching for a maximum, we maintain a probability distribution over
our \textit{estimate} of Equation~\ref{eq:lh-ratio}, and then sample
actions which are expected to improve that estimate.  This strategy
has several benefits. First, it does not make assumptions about the
scale of the similarity function. Second, by choosing to sample places
which are informative, this method implicitly minimizes the amount of
rotation.

Formally, we denote $Z_h$ as our estimate of the likelihood for
hypothesis $h$, and write its distribution as: $p(Z_h) = \int
\left[\int S(\Xb, f(\Xa, \theta, h))p(\theta)\dif\theta\right]
p(S)\dif S$, where $S$ is the similarity function, and $p(S)$ is a
prior over similarity functions.  This method of estimating an
integral is known in the machine-learning literature as
\textit{Bayesian Quadrature} \cite{Diaconis:1988uo,Osborne:2012tm}, or
BQ.

Denoting $S_h=S(\Xb, f(\Xa, \theta, h))$, we first place a
\textit{Gaussian Process} \cite{Rasmussen:2006vz}, or GP, prior on the
log of $S_h$ in order to enforce positivity after it is exponentiated,
i.e. $\mathbb{E}[Z_h] \approx \int
\exp(\mu_h(\theta))p(\theta)\dif\theta$, where $\mu_h:=\mu(\log S_h)$
is the mean of the log-GP \cite{Osborne:2012tm}.  However, this
integral is still intractable. To approximate it, we fit a second GP
over points sampled from the log-GP, which we denote as
$\bar{S}_h:=\exp(\mu_h)$. Then, from \citeA{Duvenaud:2013td}, we have
$\mathbb{E}[Z_h] \approx \int \bar{\mu}_h(\theta)p(\theta)\dif\theta$
and $\mathbb{V}(Z_h) \approx \iint \mathrm{Cov}_h(\theta,
\theta^\prime)\bar{\mu}_h(\theta)\bar{\mu}_h(\theta^\prime)p(\theta)p(\theta^\prime)\dif\theta\dif\theta^\prime$,
where $\bar{\mu}_h:=\mu(\bar{S}_h)$ is the mean of the second GP, and
$\mathrm{Cov}_h:=\mathrm{Cov}(\log S_h)$ is the covariance of the
log-GP.

Assuming independence, we can now write $p(Z_h)\approx\mathcal{N}(Z_h\
\vert\ \mathbb{E}[Z_h], \mathbb{V}(Z_h))$, which gives us a
distribution over the likelihood ratio in Equation~\ref{eq:lh-ratio}:
\begin{equation}
p(\ell)\approx\frac{\mathcal{N}(Z_0\ \vert\ \mathbb{E}[Z_0], \mathbb{V}(Z_0))\cdot{}p_0}{\mathcal{N}(Z_1\ \vert\ \mathbb{E}[Z_1], \mathbb{V}(Z_1))\cdot{}p_1}.
\end{equation}

This distribution cannot easily be calculated; however, we really are
just interested in whether $Z_0>Z_1$ or $Z_1>Z_0$. Instead, we use
$Z_D=p_0\cdot{}Z_0-p_1\cdot{}Z_1$ and compute
$p(Z_D)\propto\mathcal{N}(p_0\cdot{}\mathbb{E}[Z_0] -
p_1\cdot{}\mathbb{E}[Z_1], p_0^2\cdot{}\mathbb{V}(Z_0) +
p_1^2\cdot{}\mathbb{V}(Z_1))$.
We then sample new observations until we are at least 95\% confident
that $Z_D\neq 0$. In other words, when $p(Z_D<0)<0.025$, we accept
$\hi$, and when $p(Z_D<0)>0.975$, we accept $\hf$. Because we compare
the hypotheses in order to determine when to stop sampling, biasing
the prior towards one hypothesis should result in requiring less
evidence for that hypothesis before stopping, and more evidence for
the other hypothesis. This is in contrast to the previous models,
where we would not expect the prior to have much effect.

To determine which points to sample, we can compute the expected
variance of $Z_h$ if we sampled a new observation at $\theta_a$. From
\citeA{Osborne:2012tm}, we compute the expected variance as
$\mathbb{E}[\mathbb{V}(Z_h|\theta_a)]=\mathbb{V}(Z_h) +
\mathbb{E}[Z_h] - \int \mathbb{E}[Z_h|\theta_{a}]^2
\mathcal{N}(\mu_h(\theta_a), \mathrm{Cov}_h(\theta_a,
qc\theta_a))\dif\log S_h(\theta_a)$
for each of the actions in Equation~\ref{eq:actions}, and then choose
the action with the lowest value.

@@

We recruited 247 participants on Amazon's Mechanical Turk using the
psiTurk experiment framework \cite{Mcdonnell12}. Each participant was
paid \$1.00 for approximately 15 minutes of work, consisting of one
block of 10 practice trials followed by two blocks of 100 experiment
trials. Within a block, trials were presented in a random order.

All participants saw the same 10 practice trials as described
above. There were 720 unique experimental stimuli (20 shapes $\times$
18 angles $\times$ 2 reflections), though because stimuli with
rotations of $0^\circ$ or $180^\circ$ were repeated twice, there were
800 total experimental stimuli. These stimuli were split across eight
conditions in the following manner: first, stimuli were split into
four blocks of 200 trials. Within each block, each shape was repeated
ten times and each rotation was repeated ten times (five ``same'',
five ``flipped''), such that across all blocks, each stimulus appeared
exactly once. Each block was then split in half. Participants
completed two half-blocks (not necessarily both from the original full
block).

@@

On each trial, participants were instructed to press the `b' key to
begin and to focus on the fixation cross that appeared for 750ms
afterwards. The two images were then presented side-by-side, each at
300px $\times$ 300px, and participants could press `s' to indicate
they thought the images depicted the ``same'' object, or `d' to
indicate they thought the images depicted ``flipped'' objects.

While there was no limit on RT, participants were urged to respond as
quickly as possible while still being accurate. Specifically, we asked
them to aim for at least 85\% accuracy in the experimental blocks, and
provided a counter to keep them informed of their score.

@@

Of the 247 participants, 200 (81\%) were included in our analyses. Of
the other 47, we excluded 10 (4\%) because of an experimental error, 6
(2.4\%) because they had already completed a related experiment, and
31 (12.6\%) because they failed a comprehension check, which was
defined as correctly answering at least 85\% of stimuli with a
rotation of either $0^\circ$, $20^\circ$, or $340^\circ$.
Additionally, we excluded 82 trials for which the RT was either less
than 100ms or greater than 20s.

@@

For analyses of RT, confidence intervals around harmonic means were
computed using a bootstrap analysis of 10000 bootstrap samples
(sampled with replacement). RTs were only considered for correct
responses.  We also used a bootstrap analysis of 10000 bootstrap
samples to compute the confidence intervals around both Spearman
($\rho$) and Pearson ($r$) correlations.  Unless otherwise specified,
all correlations were computed over 720 datapoints (one for each
stimulus). For analyses of accuracy, confidence intervals were
computed from a binomial proportion with a Jeffrey's beta prior.  To
test if participants or a model was above chance on a particular
stimulus, we used the same binomial proportion and tested whether
$p\left(p(\textrm{correct})\leq 0.5\right)\leq \frac{0.05}{720}$,
where $\frac{1}{720}$ is a Bonferroni correction for multiple
comparisons.

@@

The average RT across all correctly-judged stimuli was \ExpTime{} ms;
the full histogram of RTs can be seen in
Figure~\ref{fig:histograms}. The minimum angle of rotation was
significantly rank-order (Spearman) correlated with average
per-stimulus RTs, both for ``flipped'' (\ExpThetaTimeCorrFlipped{})
and ``same'' pairs (\ExpThetaTimeCorrSame{}). While this replicates
the general result of previous experiments
\cite<e.g.,>{Shepard1971,Cooper:1975wp}, our results are not as linear
(Figure~\ref{fig:response-time-accuracy}).

The average accuracy across all stimuli was \ExpAccuracy{}, though
there were \ExpNumChance{} stimuli (out of 720) for which people were
not above chance.  The minimum angle was also correlated with
participants' average per-stimulus accuracy, though significantly more
so for ``same'' pairs (\ExpThetaAccuracyCorrSame{}) than ``flipped''
pairs (\ExpThetaAccuracyCorrFlipped{}). This is the same result found
both by \citeA{Cooper:1975wp} and \citeA{Gardony:2013gn}.

There was a significant effect of trial number both on RT
(\ExpTrialTimeCorr{}) and on accuracy (\ExpTrialAccuracyCorr{}),
though the effect on accuracy was not significant during the second
half of the experiment (\ExpaTrialAccuracyCorr{} for the first half
vs. \ExpbTrialAccuracyCorr{} for the second half). These practice
effects may have contributed to the not-quite-linearity of the human
RTs; future work will need to collect more data per participant to
mitigate this effect.

@@

The number of actions taken by the \Oc{} model was perfectly
correlated with the minimum angle of rotation
(Figure~\ref{fig:response-time-accuracy}). The \Oc{} model was the
best fit to human RTs, with a correlation of \OcTimeCorr{}
(Figure~\ref{fig:human-model-scatters}).  However, the distribution of
response times did not match that of people
(Figure~\ref{fig:histograms}). Moreover, the \Oc{} model was 100\%
accurate, and therefore could not explain the effect of rotation on
people's accuracy.

@@

There was an overall monotonic relationship between the minimum angle
of rotation and the number of actions taken by the \Th{} model
(Figure~\ref{fig:response-time-accuracy}). However, the relationship
between angle and RT for individual shapes was usually not monotonic
(e.g., Figure~\ref{fig:response-time-stimulus}).  The \Th{} model was
able to explain a moderate amount of the variance in human RTs, with a
correlation of \ThTimeCorr{}
(Figure~\ref{fig:human-model-scatters}). Like the \Oc{} model,
however, the overall distribution of its RTs did not match that of
people (Figure~\ref{fig:histograms}). Additionally, the \Th{} model
had 100\% accuracy, and thus did not exhibit a relationship between
minimum angle and accuracy.

As noted, we fit $\sigma_\epsilon=0.6$ for the \Th{} model. This has
the interesting effect of actually causing the \Th{} model to
\textit{over}rotate. This is because the step size is large enough
that it sometimes misses the global maximum, and must do another full
rotation to find it. However, when the step size is smaller, the fit
to human data is slightly worse.

@@

The \Hc{} was the only model for which there was no monotonic
relationship between rotation and RT
(Figure~\ref{fig:response-time-accuracy}). Moreover, the \Hc{} model
was barely above chance (\HcAccuracy{}) and there were \HcNumChance{}
stimuli for which it was not above chance. The \Hc{} model was not a
good predictor of human RTs (\HcTimeCorr{}), as shown in
Figure~\ref{fig:human-model-scatters}. It was a moderately good
predictor of human accuracy (\HcAccuracyCorr{}).

@@

\section{Acknowledgments} {\small This research was supported by ONR
  MURI grant number N00014-13-1-0341, and a Berkeley Fellowship
  awarded to JBH. }

>>>>

\addtolength{\abovecaptionskip}{-0.25cm}
\addtolength{\belowcaptionskip}{-0.5cm}

@@

The only models (of which the authors are aware) that seriously
attempted to address the decision of \textit{what} to simulate are
those by \citeA{Funt:1983wn} and \citeA{Just:1985uu}. In both of these
models, the axis and direction of rotation are computed prior to
performing the rotation. One object is then rotated through the target
rotation, and is checked against the other object for
congruency. However, this approach assumes that the corresponding
points on the two objects can be easily identified, which is not
necessarily the case.  Indeed, the state-of-the-art in computer vision
suggests that there is more to this problem than checking for
congruency, particularly when the shapes are complex or not exactly
the same \cite<e.g.,>{Belongie:2002tj,Sebastian:2003vm}. Additionally,
recent research shows that when performing \textit{physical}
rotations, people do not rotate until congruency is reached; they may
even rotate \textit{away} from near perfect matches
\cite{Gardony:2013gn}.

If people are not computing the rotation beforehand, what might they
be doing? To answer this question, we perform a rational analysis of
the problem of mental rotation
\cite{Marr:1983to,anderson90,Shepard:1987tt}. At the computational
level, we can say that the \textit{problem} is to determine which
spatial transformations an object has undergone based on two images of
that object (which do not include information about point
correspondences). At the algorithmic level, we are constrained by the
notion that mental images must be transformed in an analog manner (or
in a way that is approximately analog), and that mental images are
time-consuming and effortful to generate. Thus, the \textit{goal} is
to make this determination while performing a minimum amount of
computation (i.e., as few rotations as possible).

The original ``congruency'' hypothesis \cite{Shepard1971} is a
rational solution to this problem, in the sense that the smallest
amount of computation coincides with rotating through the minimum
angle. However, it violates the constraint that we do not know the
points of correspondence between the images, which is what
necessitates the use of imagery. Noting that a rational solution need
not maintain a single trajectory of rotation, we explore an
alternative model, which---rather than computing the angle of
rotation---engages in an \textit{active sampling} strategy.

Active sampling is the idea people gather new information in a manner
that increases certainty about the problem space. An everyday example
of this can be observed in the game of ``20 questions'', in which one
person thinks of a concept, and another has to guess the concept in 20
questions or less. The first question is almost always ``person,
place, or thing?'', because the answer provides the most possible
information about the concept of interest. Active sampling has
gained support across several areas of cognitive science
\cite<e.g.,>{Gureckis:2012gu}, including other spatial domains
\cite{Juni:2011vo}. In the case of mental rotation, actively choosing
rotations may be the best way to gather evidence about the similarity
between the observed shapes when the angle of rotation is unknown.

@@

The task we are interested in modeling involves observing two images
and determining whether one image depicts the ``same'' object as the
other image (differing by a rotation), or a ``flipped'' version of the
object in the other image (differing by a reflection and then a
rotation).

@@

We denote the shapes as $X_a$ and $X_b$ and assume $X_b$ is generated
by a transformation of $X_a$, i.e. $X_b=f(X_a, \theta, h)$, where
$\theta$ is a rotation, $\hi$ is the hypothesis that the images depict
the same object, and $\hf$ is the hypothesis that the images depict
mirror-image objects. The posterior probability of each hypothesis
given the observed shapes is then:
$p(h\ \vert\ X_a, X_b) \propto \int p(X_b\ \vert\ X_a, \theta,
h)p(h)p(\theta)\dif\theta$,
where $p(X_b\ \vert\ X_a, \theta, h)$ is the probability that $X_b$
was generated from $X_a$. Because we want to determine which
hypothesis is more likely, the quantity of interest is a ratio
$\ell:=p(\hi\ \vert\ X_a, X_b) / p(\hf\ \vert\ X_a, X_b)$ which
(assuming that all rotations are equally likely) is equivalent to:
\begin{equation}
  \ell = \frac{\left(\int p(X_b\ \vert\ X_a, \theta, \hi)\dif\theta\right)\cdot{}p_0}{\left(\int p(X_b\ \vert\ X_a, \theta, \hf)\dif\theta\right)\cdot{}p_1},
  \label{eq:lh-ratio}
\end{equation}
where $p_0=p(h=0)$ and $p_1=p(h=1)$, for brevity. If $\ell > 1$, then
we accept the hypothesis that the images depict the same object
($\hi$); if $\ell < 1$, then we accept the hypothesis that the images
depict flipped objects ($\hf$).

@@

We assume that the observed shapes must be transformed by a small
amount at a time, and each transformation takes a non-negligible
amount of time. If the current mental image is $\Xt$, then:
\begin{equation}
  \mathbf{X}_{t+1} = \left\{ \begin{array}{ll}
      f(\Xt, 0, \epsilon) &\mbox{ rotate by $\epsilon$ radians,} \\
      f(\Xt, 1, 0) &\mbox{ flip,} \\
      f(\Xa, 0, 0) &\mbox{ reset to $0^\circ$, or} \\
      f(\Xa, 1, 0) &\mbox{ reset and flip,} \\
    \end{array} \right.
  \label{eq:actions}
\end{equation}
where $\epsilon\sim \left|\mathcal{N}(0, \sigma_\epsilon^2)\right|$
and $\sigma_\epsilon^2$ is the variance of the step size.

@@

If we give equal weight to the two hypotheses, then the priors cancel
out; if we weigh one hypothesis more heavily, then our decision will
be biased towards that hypothesis. However, unless the likelihood
ratio is already very close to 1, small biases in the prior will not
make much of a difference.

@@

A model which does not know point correspondences could use the
following algorithm: (1) pick a random direction; (2) take a single
step; (3) if that step decreased similarity, then begin rotating in
the reverse direction, otherwise continue rotating in the original
direction; (4) continue rotating in the chosen direction until a
``match'' is found (defined as finding a value of $S$ that exceeds a
threshold); and (5) if no match was found, flip, and start over from
step one.  We only allow for the ``flip'' action after no match has
been found, because there is no particularly principled way for the
\Th{} model to choose when to flip. We assume that the locations where
$S$ is greater than the threshold correspond to the true $\theta$ (or
points near the true $\theta$). So, as with the \Oc{} model, we use
Equation~\ref{eq:mle-lh-ratio}.

@@

In the current formulation of the problem, choosing the threshold is
straightforward because we know both the exact geometry of the shapes
and that a linear transformation exists which will align
them. However, this choice is not always clear \textit{a priori}, as
the global optimum depends on many factors (e.g., shape complexity,
dimensionality, perceptual uncertainty, and whether the shapes are
identical).  One way to deal with the problem of choosing a threshold
would be use a global optimization strategy; however, this would not
result in the linear RT found by \citeA{Shepard1971}. A second
alternative is to perform a Hill Climbing (\Hc{}) search; i.e., rotate
in the direction that increases similarity until no further
improvement can be found. In contrast with the \Th{} model, this
results in arriving in a \textit{local} maximum (which may or may not
be the global maximum). Thus, as with the \Oc{} and \Th{} models, we
use Equation~\ref{eq:mle-lh-ratio}. We only allow for the ``flip''
action after a local maximum has been reached, because like the \Th{}
model, there is otherwise no principled way for the \Hc{} model to
choose when to flip.

@@

While the previous few models all focused on \textit{searching} for
the global maximum, we need only \textit{approximate}
Equation~\ref{eq:lh-ratio}. We hypothesize a model based on the idea
of \textit{active sampling} \cite<e.g.,>{Gureckis:2012gu}: instead of
searching for a maximum, we maintain a probability distribution over
our \textit{estimate} of Equation~\ref{eq:lh-ratio}, and then sample
actions which are expected to improve that estimate.  This strategy
has the benefits that it does not make assumptions about the scale of
the similarity function; and, by choosing to sample places which are
informative, this method implicitly minimizes the amount of rotation.

We denote $Z_h$ as our estimate of the likelihood for hypothesis $h$,
and write its distribution as:
$p(Z_h) = \int \left[\int S(\Xb, f(\Xa, \theta,
  h))p(\theta)\dif\theta\right] p(S)\dif S$,
where $S$ is the similarity function, and $p(S)$ is a prior over
similarity functions.  This method of estimating an integral is known
in the machine-learning literature as \textit{Bayesian Quadrature}
\cite{Diaconis:1988uo,Osborne:2012tm}, or BQ.  Denoting
$S_h=S(\Xb, f(\Xa, \theta, h))$, we first place a \textit{Gaussian
  Process} \cite{Rasmussen:2006vz}, or GP, prior on the log of $S_h$
in order to enforce positivity after it is exponentiated, i.e.
$\mathbb{E}[Z_h] \approx \int \exp(\mu_h(\theta))p(\theta)\dif\theta$,
where $\mu_h:=\mu(\log S_h)$ is the mean of the log-GP
\cite{Osborne:2012tm}.  To approximate this integral, we fit a second
GP over points sampled from the log-GP, which we denote as
$\bar{S}_h:=\exp(\mu_h)$. Then, from \citeA{Duvenaud:2013td}, we have
$\mathbb{E}[Z_h] \approx \int \bar{\mu}_h(\theta)p(\theta)\dif\theta$
and
$\mathbb{V}(Z_h) \approx \iint \mathrm{Cov}_h(\theta,
\theta^\prime)\bar{\mu}_h(\theta)\bar{\mu}_h(\theta^\prime)p(\theta)p(\theta^\prime)\dif\theta\dif\theta^\prime$,
where $\bar{\mu}_h:=\mu(\bar{S}_h)$ is the mean of the second GP, and
$\mathrm{Cov}_h:=\mathrm{Cov}(\log S_h)$ is the covariance of the
log-GP.

Assuming independence, we can now write
$p(Z_h)\approx\mathcal{N}(Z_h\ \vert\ \mathbb{E}[Z_h],
\mathbb{V}(Z_h))$,
which gives us a distribution over the likelihood ratio in
Equation~\ref{eq:lh-ratio}:
$p(\ell)\approx\mathcal{N}(Z_0\ \vert\ \mathbb{E}[Z_0],
\mathbb{V}(Z_0))\cdot{}p_0/\mathcal{N}(Z_1\ \vert\ \mathbb{E}[Z_1],
\mathbb{V}(Z_1))\cdot{}p_1$.
This distribution cannot easily be calculated, but we are only
interested in whether $Z_0>Z_1$ or $Z_1>Z_0$. So, we use
$Z_D=p_0\cdot{}Z_0-p_1\cdot{}Z_1$ and compute
$p(Z_D)\propto\mathcal{N}(p_0\cdot{}\mathbb{E}[Z_0] -
p_1\cdot{}\mathbb{E}[Z_1], p_0^2\cdot{}\mathbb{V}(Z_0) +
p_1^2\cdot{}\mathbb{V}(Z_1))$.
We then sample new observations until we are at least 95\% confident
that $Z_D\neq 0$. In other words, when $p(Z_D<0)<0.025$, we accept
$\hi$, and when $p(Z_D<0)>0.975$, we accept $\hf$. Because we compare
the hypotheses in order to determine when to stop sampling, biasing
the prior should result in requiring less evidence for one hypothesis
before stopping, and more evidence for the other hypothesis.

To choose where to sample, we compute the expected variance of $Z_h$
given a new observation at $\theta_a$. From \citeA{Osborne:2012tm}, we
compute
$\mathbb{E}[\mathbb{V}(Z_h|\theta_a)]=\mathbb{V}(Z_h) +
\mathbb{E}[Z_h] - \int \mathbb{E}[Z_h|\theta_{a}]^2
\mathcal{N}(\mu_h(\theta_a), \mathrm{Cov}_h(\theta_a,
\theta_a))\dif\log S_h(\theta_a)$
for each of the actions in Eq.~\ref{eq:actions}; we pick the one with
the lowest value.

@@

We recruited 247 participants on Amazon's Mechanical Turk using the
psiTurk experiment framework \cite{Mcdonnell12}. Each participant was
paid \$1.00 for 15 minutes of work, consisting of one block of 10
practice trials followed by two blocks of 100 randomly ordered
experiment trials.

All participants saw the same 10 practice trials as described
above. There were 720 unique experimental stimuli (20 shapes $\times$
18 angles $\times$ 2 reflections), though because stimuli with
rotations of $0^\circ$ or $180^\circ$ were repeated twice, there were
800 total experimental stimuli. These stimuli were split across eight
conditions in the following manner: first, stimuli were split into
four blocks of 200 trials. Within each block, each shape was repeated
ten times and each rotation was repeated ten times (five ``same'',
five ``flipped''), such that across all blocks, each stimulus appeared
once. Each block was then split in half, and participants completed
two half-blocks.

@@

On each trial, participants were instructed to press the `b' key to
begin and to focus on the fixation cross that appeared for 750ms
afterwards. The two images were then presented side-by-side, each at
300px $\times$ 300px, and participants could press `s' to indicate
they thought the images depicted the ``same'' object, or `d' to
indicate they thought the images depicted ``flipped'' objects.  While
there was no limit on RT, we urged participants to answer as quickly
as possible while maintaining at least 85\% accuracy in the
experimental blocks.

@@

Of the 247 participants, 200 (81\%) were included in our analyses. Of
the other 47, we excluded 10 (4\%) because of an experimental error, 6
(2.4\%) because they had already completed a related experiment, and
31 (12.6\%) because they failed a comprehension check, which was
defined as correctly answering at least 85\% of stimuli with a
rotation of either $0^\circ$, $20^\circ$, or $340^\circ$.  We also
excluded 82 trials for which the RT was either less than 100ms or
greater than 20s.

@@

For analyses of RT, confidence intervals around harmonic means of
correct responses were computed using a bootstrap analysis of 10000
bootstrap samples (sampled with replacement).  We also used a
bootstrap analysis of 10000 bootstrap samples to compute the
confidence intervals around both Spearman ($\rho$) and Pearson ($r$)
correlations.  Unless otherwise specified, all correlations were
computed over 720 stimuli. For analyses of accuracy, confidence
intervals were computed from a binomial proportion with a Jeffrey's
beta prior.  To test if judgments were above chance on a particular
stimulus, we used the same binomial proportion and tested whether
$p\left(p(\textrm{correct})\leq 0.5\right)\leq \frac{0.05}{720}$,
where $\frac{1}{720}$ is a Bonferroni correction for multiple
comparisons.

@@

The average RT across all correctly-judged stimuli was \ExpTime{} ms;
the full histogram of RTs can be seen in
Figure~\ref{fig:histograms}. The minimum angle of rotation was
significantly rank-order (Spearman) correlated with average
per-stimulus RTs, both for ``flipped'' (\ExpThetaTimeCorrFlipped{})
and ``same'' pairs (\ExpThetaTimeCorrSame{}). While this replicates
the general result of previous experiments
\cite<e.g.,>{Shepard1971,Cooper:1975wp}, our results are not as linear
(Figure~\ref{fig:response-time-accuracy}).\\
\indent The average accuracy across all stimuli was \ExpAccuracy{},
though there were \ExpNumChance{} stimuli (out of 720) for which
people were not above chance.  The minimum angle was also correlated
with participants' average per-stimulus accuracy, though much more so
for ``same'' pairs (\ExpThetaAccuracyCorrSame{}) than ``flipped''
pairs (\ExpThetaAccuracyCorrFlipped{}). This is the same
result found both by \citeA{Cooper:1975wp} and \citeA{Gardony:2013gn}.\\
\indent There was a significant effect of trial number both on RT
(\ExpTrialTimeCorr{}) and on accuracy (\ExpTrialAccuracyCorr{}),
though the effect on accuracy was not significant during the second
half of the experiment (\ExpaTrialAccuracyCorr{} for the first half
vs. \ExpbTrialAccuracyCorr{} for the second half). These effects may
have contributed to the not-quite-linearity of the human RTs; future
work should collect more data per participant.

@@

The number of actions taken by the \Oc{} model was perfectly
correlated with the minimum angle of rotation
(Figure~\ref{fig:response-time-accuracy}). The \Oc{} model was the
best fit to human RTs, with a correlation of \OcTimeCorr{}
(Figure~\ref{fig:human-model-scatters}), although the distribution of
response times did not match that of people
(Figure~\ref{fig:histograms}). Moreover, the \Oc{} model was 100\%
accurate, and therefore could not explain the effect of rotation on
people's accuracy.

@@

There was an overall monotonic relationship between the minimum angle
of rotation and the number of actions taken by the \Th{} model
(Figure~\ref{fig:response-time-accuracy}), though this relationship
did not hold for \textit{individual} shapes (e.g.,
Figure~\ref{fig:response-time-stimulus}).  The \Th{} model was able to
explain a moderate amount of the variance in human RTs, with a
correlation of \ThTimeCorr{}
(Figure~\ref{fig:human-model-scatters}). Like the \Oc{} model, the
overall distribution of its RTs did not match that of people
(Figure~\ref{fig:histograms}).  The \Th{} model had 100\% accuracy,
and thus did not exhibit a relationship between minimum angle and
accuracy.  As noted, we fit $\sigma_\epsilon=0.6$ for the \Th{}
model. This had the interesting effect of causing the \Th{} model to
\textit{over}rotate, because the step size was large enough that it
sometimes missed the global maximum, and had to do another full
rotation to find it.

@@

The \Hc{} was the only model for which there was no monotonic
relationship between rotation and RT
(Figure~\ref{fig:response-time-accuracy}). Moreover, the \Hc{} model
was barely above chance (\HcAccuracy{}) and there were \HcNumChance{}
stimuli for which it was not above chance. The \Hc{} model was not a
good predictor of human RTs (\HcTimeCorr{}), as shown in
Figure~\ref{fig:human-model-scatters}. It was a moderate predictor of
human accuracy (\HcAccuracyCorr{}).

@@

\subsubsection{Acknowledgments} {\small This research was supported by ONR
  MURI grant number N00014-13-1-0341, and a Berkeley Fellowship
  awarded to JBH. }
